### 2025-03-28 src/dataset_creation/eco_region_distribution_figure.tex
training u-net

\begin{figure}[t]
    \centering
    \includegraphics[width=\textwidth]{figures/eco_region_distribution.pdf}
    \caption{\textbf{Balanced representation of forest ecosystems across France.} 
    \textbf{a}, Geographic distribution of the 14.1 million training pixels across France's 11 ecological regions, with each pixel weighted according to the eco-region's effective forest area. 
    \textbf{b}, Comparison between the eco-region's percentage of total forest area in France (blue) and its percentage of the weighted training dataset (orange), demonstrating the high alignment achieved by our sampling strategy. 
    \textbf{c}, Relationship between pixel weights and representation ratio, showing how underrepresented regions receive higher weights (>1.0) and overrepresented regions receive lower weights (<1.0). 
    \textbf{d}, Distribution of deciduous and evergreen training pixels across ecological regions after weighting, maintaining phenological class balance within each region.}
    \label{fig:eco_region_distribution}
\end{figure}