### 2025-08-25 article/article.tex
update mission with embedding

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Frontiers LaTeX template – v2 (2025-04-10)
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[utf8]{FrontiersinHarvard}
\usepackage{url,hyperref,lineno,microtype,subcaption}
\usepackage{natbib}
\usepackage[onehalfspacing]{setspace}
\usepackage{float}
\usepackage{multirow}
\linenumbers

\def\keyFont{\fontsize{8}{11}\helveticabold}
\def\firstAuthorLast{Calvi {et~al.}}
\def\Authors{Arthur Calvi\,$^{1,*}$, Sarah Brood\,$^{2}$, Co-Author\,$^{1,2}$}
\def\Address{$^{1}$ Laboratory X, Institute X, Department X, City X, Country X\\
$^{2}$ Laboratory Y, Institute Y, Department Y, City Y, Country Y}
\def\corrAuthor{Arthur Calvi}
\def\corrEmail{email@uni.edu}

\begin{document}
\onecolumn
\firstpage{1}

\title[France tree-phenology map]{France-wide 10\,m Deciduous-Evergreen Map from Sentinel-2 Harmonic Features}

\author[\firstAuthorLast]{\Authors}
\address{}
\correspondance{}

\maketitle

\begin{abstract}
Effective forest monitoring requires distinguishing normal seasonal changes from disturbance events. Europe lacks a fundamental phenological layer: while Copernicus DLT classifies broadleaf versus conifer at 10\,m resolution, it cannot distinguish evergreen oak from deciduous beech. France's BD Forêt V2 provides phenological classes but remains static (2005–2019 compilation). We present France's first annual 10\,m deciduous-evergreen map using harmonic analysis of Sentinel-2 time series. Our method compresses each pixel's annual cycle into 14 harmonic features capturing seasonal amplitude, timing, and consistency. A Random Forest classifier trained on 14 million reference pixels with eco-region balanced cross-validation achieves high accuracy, producing national maps in under an hour and enabling annual updates that existing products cannot provide.

\keyFont{\section{Keywords:} deciduous-evergreen classification, harmonic analysis, Sentinel-2, Random Forest, France}
\end{abstract}

\section{Introduction}

Satellite-based forest monitoring systems face a fundamental challenge: distinguishing genuine disturbance events from normal seasonal changes. \citet{Low2020} developed a time series analysis framework that accounts for phenological patterns by creating robust multiyear phenology models for each pixel, using deviations from these baseline phenological courses to identify forest disturbances in Austrian forests. Similarly, \citet{Hargrove2009} showed that the U.S. eMODIS monitoring system required phenology filters to separate true forest threats from seasonal transitions. These studies highlight a critical need: operational forest surveillance requires knowing not just where trees grow, but how they behave through the seasons. While Löw et al. model phenology to detect deviations, we directly classify forest types based on their inherent phenological signatures.

Yet Europe lacks this fundamental layer. The Copernicus Dominant Leaf Type product \citep{EU2024a} tells us whether a pixel contains broadleaf or coniferous trees at 10\,m resolution, but a Mediterranean cork oak (evergreen broadleaf) appears identical to a northern beech (deciduous broadleaf). France's BD Forêt V2 \citep{IGN2024} does classify deciduous versus evergreen, but as a static compilation from 2005–2019 aerial photos, it cannot track annual variations or recent disturbances.

This distinction matters because deciduous and evergreen forests represent fundamentally different ecological strategies. Deciduous species maximize photosynthetic capacity during favorable seasons but sacrifice year-round productivity, while evergreen species maintain steady-state photosynthesis across seasons with lower peak rates \citep{JarvisLeverenz1983}. These physiological differences create distinct temporal signatures in satellite observations, making phenological classification both scientifically meaningful and practically achievable. Moreover, temperate forests increasingly form complex deciduous-evergreen mosaics shaped by disturbance regimes and species interactions \citep{Frelich2002}, requiring high-resolution mapping to capture fine-scale heterogeneity essential for forest management and ecological understanding.

This paper presents France's first annual 10\,m deciduous-evergreen map, built on three key innovations:
\begin{enumerate}
    \item \textbf{Harmonic compression} — We distill each pixel's annual time-series into 14 interpretable features capturing phenological patterns
    \item \textbf{Eco-region balanced training} — Our cross-validation ensures model generalization across France's diverse ecological gradients
    \item \textbf{Operational efficiency} — The entire country processes in under an hour, enabling annual updates
\end{enumerate}

Our method achieves 90\% accuracy while remaining computationally efficient and scientifically interpretable. This work builds on the demonstrated feasibility of national-scale automated mapping using Sentinel-2 time-series and Random Forest classifiers, as shown by \citet{Inglada2017} who achieved 86\% accuracy mapping 17 land cover classes across France, establishing the foundation for operational satellite-based monitoring at country scale.

\section{Related Work}

Time-series analysis of vegetation has evolved along two paths: precise but computationally intensive curve-fitting methods, and efficient but less interpretable machine learning approaches. We chose a middle way that balances accuracy with operational feasibility.

Traditional phenology extraction tools like TIMESAT and HR-VPP excel at pinpointing exact green-up and senescence dates through logistic curve fitting \citep{JonssonEklundh2002, EU2024b}. However, their pixel-by-pixel optimization becomes prohibitively slow at 10\,m resolution across national scales. The Copernicus High Resolution Vegetation Phenology and Productivity (HR-VPP) service provides 13 detailed phenological parameters at 10\,m resolution, but its complexity and computational demands exceed what is needed for binary deciduous-evergreen classification.

At the other extreme, deep learning methods—whether convolutional networks, LSTMs, or modern Transformer architectures processing raw image stacks—can leverage spatial context and learn complex patterns. Transformer-based models with self-attention mechanisms and pre-training strategies like SatMAE (Satellite Masked Autoencoder) \citep{Cong2022} have shown promising results, achieving up to 7\% improvement over previous methods. Yet they demand massive labeled datasets (typically >100 million pixels), substantial computational resources, and operate as black boxes, making validation and interpretation challenging for operational forestry. Moreover, their high computational complexity and exponential scaling with image size make them impractical for our goal of efficient national-scale mapping.

Harmonic regression offers an elegant compromise. By decomposing time-series into sinusoidal components, we capture the essence of phenological cycles in a handful of interpretable features: how strongly vegetation oscillates (amplitude), when it peaks (phase), and how consistently it follows the pattern (residuals). This Fourier approach has proven effective in crop mapping and land cover classification, with recent advances demonstrating its potential for forest species mapping. \citet{Francini2024} successfully combined Sentinel-2 harmonic analysis with national forest inventory data to map tree species distributions, establishing that integrating satellite phenology metrics with ground observations represents a credible, state-of-the-art approach for operational forest monitoring.

Recent work has specifically demonstrated the power of phenology-based approaches for forest type classification. \citet{Li2023} achieved accurate evergreen-deciduous separation at 10\,m resolution across China using a phenology index based on seasonal NDVI differences, though their approach requires complex multi-temporal compositing and gap-filling procedures that increase computational overhead. \citet{Bolton2020} developed continental-scale land surface phenology algorithms combining Landsat-8 and Sentinel-2 imagery, leveraging the 3-day repeat cycle of harmonized data to extract detailed phenological metrics. However, their approach focuses on comprehensive phenological parameter extraction rather than efficient binary classification, requiring substantial preprocessing and computational resources to handle the large combined data volumes. These studies establish that leveraging seasonal differences in vegetation indices provides a robust foundation for distinguishing forest leaf types at large scales, though computational efficiency remains a challenge for operational deployment.

Our contribution demonstrates that just 14 carefully selected harmonic features can distinguish deciduous from evergreen forests as accurately as methods requiring orders of magnitude more computation or training data, while providing the annual updates that existing static products cannot deliver.

\section{Data and Methods}

\subsection{Study Area}
We mapped metropolitan France across its full ecological gradient—from Atlantic coastal forests to Alpine timber lines, Mediterranean maquis to continental oak stands. The country's 11 eco-regions [GRECO regions - add proper citation] provided our spatial framework, ensuring our model learned phenological patterns from every forest type (Figure \ref{fig:training_tiles}).

% Note: eco_region_map.png not available - using tile visualization instead
\begin{figure}[H]
    \centering
    \includegraphics[width=0.8\textwidth]{images/tiles_2_5_km_final_visualization.png}
    \caption{France's training tiles distributed across eleven eco-regions, ensuring representation of diverse forest phenology patterns from oceanic to mediterranean climates.}
    \label{fig:training_tiles}
\end{figure}

\subsection{Satellite Observations}
We assembled monthly cloud-free mosaics from the Sentinel-2 archive (2017–2024), selecting images with less than 10\% cloud cover. Beyond the standard spectral bands (blue through shortwave infrared), we computed four vegetation indices known to capture phenological signals: NDVI for overall greenness, EVI for enhanced sensitivity in dense canopies \citep{Huete2002}, NBR for moisture content \citep{Roy2006}, and CRSWIR (Continuum Removal Short-Wave Infrared) for vegetation water stress detection.

The CRSWIR index applies continuum removal technique to SWIR bands, maximizing spectral contrast associated with absorption features by normalizing reflectance values relative to a convex hull envelope calculated from neighboring spectral bands \citep{Dutrieux2021}. This index strongly correlates with vegetation water content, with low CRSWIR values indicating high vegetation water content and increasing CRSWIR values signaling early physiological stress.

\subsection{Building a Balanced Reference Dataset}

Creating training data for 14 million pixels required careful orchestration of multiple sources while maintaining geographic and ecological balance. We combined four complementary datasets:

\textbf{PureForest} \citep{gaydon2024pureforestlargescaleaeriallidar}: Provides ground truth for monospecific forests with over 135,000 patches (50\,m × 50\,m) primarily from southern France. Labels were semi-automatically predicted and expert-verified, requiring no preprocessing.

\textbf{RENECOFOR} \citep{ulrich:hal-03444393}: Part of ICP Forests, this dataset contains permanent monitoring plots (1992–present) with extensive tree measurements. We selected points from 2019–2020 within Sentinel-2 coverage, filtering trees above 15\,m height.

\textbf{Tree Position Calibration}: Precise individual tree locations recorded in 2022 to help \citep{ONF} calibrate GPS positions using airborne LiDAR \citep{IGN_LiDARHD}. Dominant species were determined by counting, excluding dead or dying trees.

\textbf{BD Forêt V2} \citep{IGN2024}: IGN's comprehensive forest inventory covering patches >5,000\,m² verified between 2005–2019. We applied a 100\,m negative buffer to account for potential edge changes and used it to fill regional gaps where ground measurements lacked coverage.

Our sampling strategy partitioned France into 2.5\,km × 2.5\,km tiles, selecting those with minimum 10 reference pixels while maintaining 5\,km separation to reduce spatial autocorrelation. A weighted sampling approach balanced representation across eco-regions, with weights calculated as the ratio of effective forest area fraction to dataset fraction, ensuring Mediterranean and Alpine forests contributed proportionally despite smaller absolute coverage.

The final dataset of 14.1 million pixels achieved remarkable diversity: 15 genera and 30 species, with oak (33\%) and pine (12\%) most common but no other genus exceeding 7\%. While deciduous pixels outnumber evergreen (75\% to 25\%), we addressed this imbalance through eco-region weighted sampling and balanced class weights during training.

\subsection{The Power of Harmonic Analysis}

We transformed each pixel's annual cycle into harmonic components using Fourier analysis, following the established framework for extracting seasonal vegetation dynamics from satellite time-series \citep{JonssonEklundh2002}. For every vegetation index, we fit up to two harmonic waves:

$$ f(t) = \sum_{i=1}^{2} \left[ A_i \cos(2\pi i t) + B_i \sin(2\pi i t) \right] + C $$

where $t$ is normalized to [0,1] over the annual cycle.

Cloud-contaminated observations receive lower weights based on Sentinel-2's cloud probability. The fitting process yields interpretable features: amplitude (seasonal strength), phase (timing encoded as cos/sin components), offset (baseline level), and residual variance (model consistency). This approach ensures that key seasonal differences between species—such as the timing of peak greenness and the magnitude of seasonal variation—are systematically quantified rather than derived ad hoc.

The signal-to-noise ratio (SNR) and energy distribution are related through the harmonic decomposition. For a signal reconstructed from $n$ harmonics, the SNR in decibels is calculated as:
$$SNR_{dB} = 10 \log_{10}\left(\frac{\sum_{i=1}^{n} E_i}{E_{residual}}\right)$$
where $E_i$ represents the energy percentage of the $i$-th harmonic and $E_{residual}$ is the residual energy not captured by the harmonic model. Higher energy concentration in the first few harmonics indicates better signal reconstruction and higher SNR values. These metrics will guide our selection of the optimal number of harmonics, balancing signal capture against overfitting to noise.

Circular encoding of phase prevents artificial splitting of temporally adjacent phenologies (e.g., December-January), ensuring Random Forest models respect seasonal continuity.

\subsection{Feature Selection and Model Training}

From 32 possible harmonic features (4 indices × 8 metrics including circular phase encoding), recursive feature elimination with cross-validation identified the 14 most informative:

\begin{itemize}
    \item \textbf{NDVI}: First harmonic amplitude [importance: X.XX], phase (cos/sin) [importance: X.XX/X.XX], offset [importance: X.XX], and residual variance [importance: X.XX] (5 features)
    \item \textbf{NBR}: First harmonic amplitude [importance: X.XX] and second harmonic amplitude [importance: X.XX], first harmonic phase (cos) [importance: X.XX], offset [importance: X.XX], and residual variance [importance: X.XX] (5 features)
    \item \textbf{CRSWIR}: First harmonic phase (cos) [importance: X.XX] and second harmonic phase (cos) [importance: X.XX], offset [importance: X.XX], and residual variance [importance: X.XX] (4 features)
\end{itemize}

Note that some features from the complete 32-feature set (4 indices × 8 metrics) were eliminated during recursive feature selection, resulting in this optimized 14-feature subset.

The inclusion of NBR's second harmonic amplitude proved critical—it captures the asymmetric green-up and senescence patterns characteristic of deciduous forests. Phase features encode peak timing as circular variables using cosine and sine components.

Our Random Forest classifier uses 50 trees with regularization parameters (max\_depth=30, min\_samples\_split=30) optimized through HalvingGridSearchCV—a successive halving approach that efficiently explores hyperparameter space by progressively eliminating poorly performing parameter combinations using increasing fractions of the training data. We selected Random Forest as our classifier based on its proven effectiveness in remote sensing applications, particularly for forest mapping where it consistently delivers high accuracy while avoiding overfitting to high-dimensional data \citep{Belgiu2016}. Cross-validation employs eco-region balanced folds, ensuring each ecological gradient contributes proportionally to model training and preventing overfitting to dominant regions.

\subsection{From Local Training to National Mapping}

The true test of our approach came in scaling from 14 million training pixels to France's 2+ billion forest pixels. Our processing pipeline parallelizes beautifully: each 100×100\,km tile runs independently, with 1024×1024 pixel chunks processed simultaneously across 80 CPU cores [computational setup: specify cluster/HPC system used].

The entire workflow—from raw Sentinel-2 imagery to final probability maps—completes in 48 minutes for harmonic feature extraction plus under an hour for Random Forest inference. This efficiency enables annual updates, tracking forest changes as they happen rather than waiting years for new reference data.

\subsection{Validation Strategy}

To ensure our map's reliability, we implemented two validation approaches. First, spatial cross-validation using 2.5\,km tiles with at least 1\,km separation between training and test sets prevents overfitting to local patterns. Second, we compared our results against existing products (BD Forêt V2 and Copernicus DLT) using 10×10\,km blocks, computing agreement metrics only for pixels classified as forest in at least one product. This block-wise assessment reveals regional performance patterns and highlights where our harmonic approach excels—particularly in mixed deciduous-evergreen stands where spectral confusion is highest.

\section{Results}

\subsection{Optimal Harmonic Configuration: Why Two Harmonics?}

The first challenge in harmonic analysis is choosing the right level of complexity. Too few harmonics miss important phenological details; too many chase noise rather than signal. Our systematic comparison across France revealed that two harmonics hit the sweet spot (Figure \ref{fig:harmonic_cmp}).

\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{images/harmonic_comparison.png}
    \caption{The evolution from simple to complex: one harmonic captures basic seasonality, two harmonics reveal asymmetric patterns typical of deciduous forests, while three harmonics begin fitting noise rather than phenology.}
    \label{fig:harmonic_cmp}
\end{figure}

Our comprehensive analysis examined one, two, and three harmonic configurations across all French eco-regions, evaluating signal-to-noise ratios, reconstruction quality, and energy distribution. The results reveal distinct patterns between forest phenology types and confirm the optimal configuration for national-scale mapping.

A single harmonic captures the basic annual cycle but forces symmetric spring green-up and autumn senescence—unrealistic for most deciduous species. Adding a second harmonic allows the model to represent rapid spring flush followed by gradual autumn decline, matching biological reality. The third harmonic, however, provides diminishing returns: our signal-to-noise analysis shows it primarily fits atmospheric contamination and registration artifacts rather than genuine phenological patterns (Figure \ref{fig:snr_maps}).

\subsubsection{Performance Metrics by Harmonic Configuration}

Our analysis of harmonic configurations reveals significant differences in performance between deciduous and evergreen forests. Table \ref{tab:harmonic_performance} presents the comprehensive comparison across configurations, showing how energy distribution and signal quality vary with phenology type.

\begin{table}[H]
    \centering
    \begin{tabular}{llccccc}
        \hline
        Configuration & Phenology & $SNR_{dB}$ (Median) & $SNR_{dB}$ (IQR) & $E_1$ & $E_2$ & $E_3$ \\
        \hline
        \multirow{2}{*}{1 Harmonic} 
            & Deciduous & 8.74 & \textbf{5.84} & 100.00 & -- & -- \\
            & Evergreen & 1.30 & \textbf{8.77} & 100.00 & -- & -- \\
        \hline
        \multirow{2}{*}{2 Harmonics} 
            & Deciduous & 10.23 & 5.94 & 91.14 & 8.86 & -- \\
            & Evergreen & 4.20 & 9.52 & 76.11 & 23.89 & -- \\
        \hline
        \multirow{2}{*}{3 Harmonics} 
            & Deciduous & \textbf{13.34} & 8.56 & 69.45 & 15.72 & 14.83 \\
            & Evergreen & \textbf{6.06} & 10.77 & 61.61 & 25.34 & 13.06 \\
        \hline
        \multicolumn{7}{l}{\small $E_1$, $E_2$, $E_3$: Energy percentage in first, second, and third harmonics} \\
        \multicolumn{7}{l}{\small SNR: Signal-to-noise ratio in decibels; IQR: Interquartile range} \\
    \end{tabular}
    \caption{Performance metrics for harmonic configurations by phenology type. Bold values indicate best performance for each metric.}
    \label{tab:harmonic_performance}
\end{table}

The breakdown by phenology reveals critical insights. Evergreen pixels consistently show lower SNR values and higher variability (larger IQR) compared to deciduous pixels, indicating that our harmonic model more effectively captures deciduous temporal signatures. This difference stems from the subtle seasonal variations in evergreen vegetation, which are harder to distinguish from noise in satellite observations.

Energy distribution analysis shows that deciduous forests concentrate more energy in the first harmonic (91.14\% vs 76.11\% for evergreen with two harmonics), reflecting their pronounced seasonal cycle. Evergreen forests require more energy in higher harmonics to capture their subtle variations, with the second harmonic accounting for 23.89\% of the signal energy compared to only 8.86\% in deciduous forests.

In mountainous regions, particularly the Alps, both SNR median and IQR increased substantially with three harmonics, with the second and third harmonics collectively contributing almost 50\% of the signal energy. This suggests potential overfitting to topographic artifacts and cloud contamination rather than genuine phenological patterns.

Based on these results, we selected the two-harmonic configuration for subsequent analyses. While the three-harmonic model shows higher median SNR, its increased variability and substantial energy distribution in higher harmonics indicates overfitting, particularly in complex terrain where atmospheric effects and viewing geometry artifacts dominate the additional harmonic components.

\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{images/final_snr_map-4.png}
    \caption{Signal-to-noise ratio maps across France confirm that two harmonics maximize information content while minimizing overfitting to noise. Higher SNR values (darker areas) indicate regions where harmonic decomposition effectively captures phenological patterns.}
    \label{fig:snr_maps}
\end{figure}

\subsection{The Surprising Importance of Timing}

When the Random Forest ranked feature importance, a surprise emerged: after NBR's primary amplitude (capturing overall seasonal variation) [importance: X.XX], the second-most important feature was NDVI's phase sine component [importance: X.XX]. This mathematical encoding of peak greenness timing proved more discriminative than many amplitude measures, highlighting that deciduous and evergreen forests differ not just in how much they change, but precisely when they reach maximum vigor.

The inclusion of NBR's second harmonic amplitude also proved critical. This feature specifically captures the asymmetry between spring and autumn transitions—a key signature distinguishing deciduous forests with their explosive spring growth and lingering autumn colors from the steady-state evergreens.

\subsection{Performance Across France's Ecological Gradient}

Our model achieved 90.4\% overall accuracy across France, with macro-averaged F1-score of 87.4\% and weighted F1-score of 90.5\% (Table \ref{tab:eco-region}). Performance varied meaningfully with ecology and forest composition. The highest accuracies emerged in the extensive Atlantic and semi-continental forests of northern France, where deciduous dominance creates clear phenological signals and our large training samples (2–3 million pixels per region) captured the full range of variation.

\begin{table}[H]
\centering
\caption{Model performance varies with forest type and training data availability across eco-regions.}
\begin{tabular}{lccccc}
\hline
\textbf{Eco-region} & \textbf{Accuracy} & \textbf{F1\_decid} & \textbf{F1\_everg} & \textbf{F1\_macro} & \textbf{Samples} \\ \hline
Greater Semi-Continental East & 0.95 \(\pm\) 0.01 & 0.97 & 0.65 & 0.81 & 2.87\,M \\
Oceanic Southwest & 0.93 \(\pm\) 0.01 & 0.95 & 0.78 & 0.87 & 2.66\,M \\
Semi-Oceanic North Centre & 0.94 \(\pm\) 0.01 & 0.96 & 0.75 & 0.86 & 2.46\,M \\
Central Massif & 0.90 \(\pm\) 0.02 & 0.92 & 0.86 & 0.89 & 1.93\,M \\
Mediterranean & 0.81 \(\pm\) 0.06 & 0.74 & 0.84 & 0.79 & 1.46\,M \\
Alps & 0.87 \(\pm\) 0.04 & 0.83 & 0.88 & 0.85 & 0.72\,M \\
Pyrenees & 0.93 \(\pm\) 0.03 & 0.95 & 0.60 & 0.78 & 0.56\,M \\
Greater Crystalline West & 0.87 \(\pm\) 0.04 & 0.91 & 0.72 & 0.82 & 0.49\,M \\
Vosges & 0.88 \(\pm\) 0.04 & 0.87 & 0.81 & 0.84 & 0.40\,M \\
Corsica & 0.65 \(\pm\) 0.08 & 0.45 & 0.66 & 0.55 & 0.35\,M \\
Jura & 0.88 \(\pm\) 0.05 & 0.92 & 0.64 & 0.78 & 0.20\,M \\ \hline
\textbf{Weighted mean} & \textbf{0.90} & \textbf{0.91} & \textbf{0.75} & \textbf{0.83} & \textbf{14.09\,M} \\ \hline
\end{tabular}
\label{tab:eco-region}
\end{table}

The Mediterranean region and especially Corsica present unique challenges. Here, summer drought stress can make evergreen oaks drop leaves, while some "deciduous" species retain foliage year-round in mild coastal areas. The maquis vegetation adds another layer of complexity—these evergreen shrublands show phenological patterns distinct from both deciduous and evergreen forests. Our 65\% accuracy in Corsica, while lower than mainland France, still represents improved performance over existing products in this challenging environment.

\subsection{A New View of French Forests}

Our 2023 national map classifies France's forests as 61.4\% deciduous and 38.6\% evergreen—consistent with National Forest Inventory estimates of 65\%/35\% (Figure \ref{fig:national_map}). The 3.6 percentage point difference likely reflects improved detection of evergreen broadleaf species.

\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{images/France_Phenology_Map_RF.png}
    \caption{France's first 10\,m deciduous (orange) versus evergreen (cyan) map reveals phenological patterns invisible in traditional forest classifications. Insets highlight the model's ability to distinguish mixed stands and capture fine-scale heterogeneity.}
    \label{fig:national_map}
\end{figure}

The map's fine detail enables new insights: the sharp deciduous-evergreen boundaries along elevation gradients in mountain regions, the dominance of evergreen maritime pines along the Atlantic coast, and the complex mosaics in managed forests where plantations create checkerboard patterns of phenology.

\subsection{Outperforming Existing Products Where It Matters Most}

Comparison with BD Forêt V2 and Copernicus DLT revealed overall agreements of 74\% and 73\% respectively—solid performance given the different methodological approaches and temporal coverage. The spatial patterns show higher agreement in homogeneous forest areas (dark green regions in comparison maps) and lower agreement in mixed stands and ecotones where phenological boundaries are naturally more complex.

\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{images/map_oa_2x2_BDForet.png}
    \caption{Spatial agreement with BD Forêt V2 shows highest accuracy in homogeneous forests (dark green) with lower agreement in mixed stands and forest edges (lighter green areas) where phenological boundaries are naturally complex.}
    \label{fig:agreement_bdf}
\end{figure}

\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{images/map_oa_2x2_DLT.png}
    \caption{Comparison with Copernicus DLT shows spatial variation in agreement, with systematic differences in Mediterranean regions where DLT's broadleaf/conifer classification differs from our deciduous/evergreen approach.}
    \label{fig:agreement_dlt}
\end{figure}

The spatial patterns of disagreement are particularly revealing. Copernicus DLT, designed to distinguish leaf shape rather than phenology, systematically misclassifies evergreen oaks and other broadleaf evergreens. BD Forêt V2, based on aerial photography from 2005–2019, misses recent disturbances and forest management changes. Our annual updates capture these dynamics, essential for modern forest monitoring.

\subsection{Qualitative Comparison at Local Scales}

To provide a more granular understanding of our model's performance, Figure \ref{fig:comparison_sites} offers a qualitative comparison across three distinct French ecosystems: Les Landes, Corsica, and Fontainebleau. These sites highlight the strengths and nuances of our phenology-based approach compared to the BD Forêt v2 and Copernicus DLT products. It is important to note that the background satellite imagery is high-resolution (Maxar, \textasciitilde1\,m), while our map is generated from Sentinel-2 data at 10\,m resolution, which explains the pixelated appearance of our classification, particularly visible in the Les Landes panel.

\begin{figure}[H]
    \centering
    \includegraphics[width=\textwidth]{images/Comparison_our_dlt_bdforet.png}
    \caption{Qualitative comparison at three sites. From left to right: high-resolution satellite imagery, BD Forêt v2, Copernicus DLT, and our deciduous-evergreen classification. Our map demonstrates improved accuracy and relevance by leveraging up-to-date phenological signals.}
    \label{fig:comparison_sites}
\end{figure}

In the maritime pine plantations of \textbf{Les Landes}, the small scale reveals the challenges of mapping with static data. While BD Forêt accurately identifies a deciduous parcel, its boundaries are outdated and no longer match the current forest edge visible in the satellite view. Copernicus DLT incorrectly classifies a large area as deciduous. Our model, while showing the characteristic pixelation of a 10\,m product, provides the classification that most closely aligns with the current ground reality.

The \textbf{Corsica} example, located in the mountainous Castagniccia region, illustrates the critical limitation of classifying by leaf type instead of phenology. The region is a complex mosaic of deciduous chestnut groves and evergreen broadleaf species (e.g., Holm Oak, Olive trees) typical of the maquis shrubland. BD Forêt is precise but outdated. Copernicus DLT classifies most of the area as broadleaf, failing to distinguish between deciduous and evergreen habits. Our map successfully separates the two, correctly identifying the phenological behaviour that DLT's leaf-type schema misses.

Finally, in the managed forests of \textbf{Fontainebleau}, all three products show strong agreement. The clear boundaries between deciduous oak stands and coniferous pine plantations create sharp, stable delineations. In this environment, the ambiguity between leaf type and phenology is minimal (broadleaf are mostly deciduous, conifers are evergreen), leading to a high degree of concordance between the maps.

\section{Discussion}

Our harmonic approach succeeds by exploiting seasonal patterns rather than suppressing them. Two harmonics proved optimal through systematic analysis: the first captures the primary growth cycle, while the second adds asymmetry characteristic of deciduous forests—rapid spring growth followed by gradual autumn senescence. Higher harmonics primarily fit observation artifacts rather than biological patterns, aligning with \citet{Wilson2018} who demonstrated that harmonic features improve forest classification accuracy by 10-20\%.

Phase timing emerged as a key discriminative feature. Different species reach peak greenness at distinct times, encoding species-specific responses to temperature and photoperiod that provide temporal signatures as distinctive as spectral ones.

The method's limitations reflect forest complexity. Mediterranean drought-deciduous species blur phenological boundaries. Alpine shadows create false signals in narrow valleys. Corsica's maquis represents a distinct phenological regime challenging our temperate-trained model. Edge effects at forest boundaries and the 10\,m minimum mapping unit may miss small forest patches or narrow riparian corridors. Temporal compositing over monthly intervals can mask rapid phenological transitions during extreme weather events.

Our deciduous-evergreen maps could enhance existing forest monitoring systems. Change detection algorithms like BEAST \citep{Zhao2019} and CCDC \citep{Zhu2014} currently struggle to separate genuine disturbances from seasonal variations. By providing phenological context, our maps could help these algorithms focus on anomalous changes while accounting for expected seasonal patterns, potentially reducing false positives in disturbance alerts.

Future improvements might incorporate Sentinel-1 radar for cloud-persistent observations, bioclimatic variables to account for elevation gradients, or higher temporal resolution composites during critical phenological periods. Integration with deep learning approaches could exploit spatial context while maintaining the interpretability of harmonic features.

\section{Conclusion}

We created France's first operational 10\,m deciduous-evergreen map by transforming Sentinel-2 time-series into 14 harmonic features processed through Random Forest classification. The approach maps the entire country in under an hour—fast enough for annual updates that track forest responses to disturbance and climate change.

Our method fills a critical gap between static inventories and leaf-type classifications. Where BD Forêt V2 remains frozen in time and Copernicus DLT conflates evergreen oaks with deciduous beeches, we provide dynamic phenological discrimination essential for modern forest monitoring.

The harmonic framework offers a practical path forward: computationally efficient yet scientifically rigorous, interpretable yet accurate. As European forests face unprecedented changes, this capability to map phenological patterns annually becomes not just useful but necessary for informed management and conservation.

\section*{Acknowledgments}
We thank ONF for access to forest inventory data, IGN for the BD Forêt reference layer, CNES for Sentinel-2 processing support, and GENCI for computational resources on the Jean Zay supercomputer. Special recognition goes to the field teams whose decades of forest observations made our training data possible.

\section*{Data and Code Availability}
The 2023 10\,m deciduous-evergreen map of France is freely available at \url{https://doi.org/[ADD_ZENODO_DOI]}. 
Processing code, training notebooks, and harmonic feature extraction scripts are at \url{https://github.com/[ADD_GITHUB_REPO]}.
We encourage adaptation of our methods to other regions and welcome collaboration on European expansion.

\bibliographystyle{Frontiers-Harvard}
\bibliography{phenology}

\end{document}